The Sound of Teaching Music: Expert Pianists' Performance Modulations for Novices

### 2020-12-15 paper/manuscript.Rmd
add some figs
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

@@

We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment. For data analysis, we excluded 5 participants due to experimental errors (3), an overall deviated tempo (2). Thirty-one participants (`r nrow(dt_included[Gender == "Female"])` female) were included in data analysis. Most participants were right-handed (left; `r nrow(dt_included[Handedness == "Left"])`, both; `r nrow(dt_included[Handedness == "Both"])`). They had `r round(mean(dt_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_included$PianoTotalPractice), 2)`). All participants gave their informed consent before the experiment started and received vouchers for their participation. The study was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary.

@@

# Combine
img_stimuli <- image_append(c(stim_n, stim_a, stim_d), stack = TRUE)
img_stimuli <- image_annotate(img_stimuli, "(A)", size = 80, location = "+100+0", color = "black", font = "helvetica")
img_stimuli <- image_annotate(img_stimuli, "(B)", size = 80, location = "+100+650", color = "black", font = "helvetica")
img_stimuli <- image_annotate(img_stimuli, "(C)", size = 80, location = "+100+1300", color = "black", font = "helvetica")
img_stimuli
```

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs. performing) and Subcomponent-LS (legato vs. staccato) revealed that there was a significant main effect of Subcomponent-LS (F) and a significant interaction between Condition and Subcomponent-LS (F). Post-hoc comparisons showed that participants produced shorter staccato in the teaching condition than in the performing condition (t) whereas there was no significant difference in legato between the teaching and performing condition (t).

We also calculated KOR...

@@

@@

### 2.3.2. Key Velocity (KV)
### 2.3.3. KV Transition Points
### 2.3.4. Key-Overlap Time (KOT)

@@

@@

# Combine
img_stimuli_2 <- image_append(c(stim_n_2, stim_a_2, stim_d_2), stack = TRUE)
img_stimuli_2 <- image_annotate(img_stimuli_2, "(A)", size = 80, location = "+100+70", color = "black", font = "helvetica")
img_stimuli_2 <- image_annotate(img_stimuli_2, "(B)", size = 80, location = "+100+1125", color = "black", font = "helvetica")
img_stimuli_2 <- image_annotate(img_stimuli_2, "(C)", size = 80, location = "+100+2150", color = "black", font = "helvetica")
img_stimuli_2
```

@@

### 3.2.2. Key-Overlap Time (KOT)
### 3.2.3. Key Velocity (KV)
### 3.2.4. KV Transition Points

@@

### 3.3.2. Key Velocity (KV)
### 3.3.3. KV Transition Points
### 3.3.4. Key-Overlap Time (KOT)

@@

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

>>>>

abstract: |
  aaaa

@@

We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment. For data analysis, we excluded 5 participants due to experimental errors (3), an overall deviated tempo (2; outside 2 standard deviations from the average tempo across participants). Thirty-one participants (`r nrow(dt_included[Gender == "Female"])` female) were included in data analysis. Most participants were right-handed (left; `r nrow(dt_included[Handedness == "Left"])`, both; `r nrow(dt_included[Handedness == "Both"])`). They had `r round(mean(dt_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_included$PianoTotalPractice), 2)`). All participants gave their informed consent before the experiment started and received vouchers for their participation. The study was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary.

@@

# combine
img_stimuli <- image_append(c(stim_n, stim_a, stim_d), stack = TRUE)
img_stimuli <- image_annotate(img_stimuli, "(A)", size = 80, location = "+100+0", color = "black", font = "helvetica")
img_stimuli <- image_annotate(img_stimuli, "(B)", size = 80, location = "+100+650", color = "black", font = "helvetica")
img_stimuli <- image_annotate(img_stimuli, "(C)", size = 80, location = "+100+1300", color = "black", font = "helvetica")
img_stimuli
```

@@

Blah

```{r kot-art-1, include = FALSE}
kot_art_plot <- data.table(subject = kot_art[Condition == "teaching"]$SubNr, teaching = kot_art[Condition == "teaching"]$Mean, performing = kot_art[Condition == "performing"]$Mean, subcomponent = kot_art[Condition == "teaching"]$Subcomponent)

kot_art_1 <- ggpaired(kot_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation") +
  ylim(-170, 105) +
  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = -50, paired = TRUE) +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "kot_art_1.png"), plot = kot_art_1, dpi = 600, width = 7, height = 5)
img_kot_art_1 <- image_read(here("paper/image", "kot_art_1.png"))
```

@@

Blah

@@

### 2.3.2. Key Velocity (KV)

Blah

### 2.3.3. KV Transition Points

Blah

### 2.3.4. Key-Overlap Time (KOT)

Blah

```{r kot-dyn-1, include = FALSE}
kot_dyn_plot <- data.table(subject = kot_dyn[Condition == "teaching"]$SubNr, teaching = kot_dyn[Condition == "teaching"]$Mean, performing = kot_dyn[Condition == "performing"]$Mean, subcomponent = kot_dyn[Condition == "teaching"]$Subcomponent)

kot_dyn_1 <- ggpaired(kot_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics") +
  ylim(-170, 105) +
  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 0, paired = TRUE) +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "kot_dyn_1.png"), plot = kot_dyn_1, dpi = 600, width = 7, height = 5)
img_kot_dyn_1 <- image_read(here("paper/image", "kot_dyn_1.png"))
```

@@

```{r plot-kot-1, out.width = "100%", fig.cap = "\\label{fig:kot-1}Experiment 1: KOT (ms) when playing the piece with either articulation (left) or dynamics (right). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box. Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
# combine
img_kot_1 <- image_append(c(img_kot_art_1, img_kot_dyn_1))
image_write(img_kot_1, path = here("paper/image", "img_kot_1.png"), format = "png")
img_kot_1
```

@@

# combine
img_stimuli_2 <- image_append(c(stim_n_2, stim_a_2, stim_d_2), stack = TRUE)
img_stimuli_2 <- image_annotate(img_stimuli_2, "(A)", size = 80, location = "+100+70", color = "black", font = "helvetica")
img_stimuli_2 <- image_annotate(img_stimuli_2, "(B)", size = 80, location = "+100+1125", color = "black", font = "helvetica")
img_stimuli_2 <- image_annotate(img_stimuli_2, "(C)", size = 80, location = "+100+2150", color = "black", font = "helvetica")
img_stimuli_2
```

@@

### 3.2.2. Key-Overlap Ratio (KOR)

Blah

```{r kot-art-2, include = FALSE}
kot_art_plot <- data.table(subject = kot_art[Condition == "teaching"]$SubNr, teaching = kot_art[Condition == "teaching"]$Mean, performing = kot_art[Condition == "performing"]$Mean, subcomponent = kot_art[Condition == "teaching"]$Subcomponent)

kot_art_2 <- ggpaired(kot_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
  ylim(-1, 0.4) +
  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = -0.5, paired = TRUE) +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "kot_art_2.png"), plot = kot_art_1, dpi = 600, width = 7, height = 5)
img_kot_art_2 <- image_read(here("paper/image", "kot_art_2.png"))
```

### 3.2.3. Key Velocity (KV)

Blah

### 3.2.4. KV Transition Points

Blah

@@

### 3.3.2. Key Velocity (KV)

Blah

### 3.3.3. KV Transition Points

Blah

### 3.3.4. Key-Overlap Ratio (KOR)

Blah

```{r kot-dyn-2, include = FALSE}
kot_dyn_plot <- data.table(subject = kot_dyn[Condition == "teaching"]$SubNr, teaching = kot_dyn[Condition == "teaching"]$Mean, performing = kot_dyn[Condition == "performing"]$Mean, subcomponent = kot_dyn[Condition == "teaching"]$Subcomponent)

kot_dyn_2 <- ggpaired(kot_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
  ylim(-1, 0.4) +
  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = -0.5, paired = TRUE) +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "kot_dyn_2.png"), plot = kot_dyn_1, dpi = 600, width = 7, height = 5)
img_kot_dyn_2 <- image_read(here("paper/image", "kot_dyn_2.png"))
```

@@

\begingroup
\setlength{\parindent}{-0in}
\setlength{\leftskip}{0in}

### 2020-12-15 paper/manuscript.tex
add some figs
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

\abstract{
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general context}.

Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
}

@@

We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment. For data analysis, we excluded 5 participants due to experimental errors (3), an overall deviated tempo (2). Thirty-one participants (15 female) were included in data analysis. Most participants were right-handed (left; 2, both; 2). They had 12.45 years of practice on average (\emph{SD} = 5.66). All participants gave their informed consent before the experiment started and received vouchers for their participation. The study was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary.

@@

In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution. The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.21{]} while playing the piece with notated articulation (\emph{p} = 0.017; \emph{Figure \ref{fig:ioi-1}}).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (F) and a significant interaction between Condition and Subcomponent-LS (F). Post-hoc comparisons showed that participants produced shorter staccato in the teaching condition than in the performing condition (t) whereas there was no significant difference in legato between the teaching and performing condition (t).

We also calculated KOR\ldots{}

@@

The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.68 (ms), \emph{IQR} = 21.99{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 20.04{]} while playing the piece with notated dynamics (\emph{p} = 0.120; \emph{Figure \ref{fig:ioi-1}}).

@@

@@

A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.966, \emph{SD} = 0.049{]} than in the performing condition {[}\emph{M} = 0.953, \emph{SD} = 0.040{]} while playing the piece with notated articulation (\emph{t}(19) = 2.47
, \emph{p} = 0.023; \emph{Figure \ref{fig:ioi-2}}). A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (\emph{p} = 0.033).

\hypertarget{key-overlap-time-kot-2}{%
\subsubsection{3.2.2. Key-Overlap Time (KOT)}\label{key-overlap-time-kot-2}}

@@

A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.957, \emph{SD} = 0.048{]} and the performing condition {[}\emph{M} = 0.956, \emph{SD} = 0.043{]} while playing the piece with notated articulation (\emph{t}(19) = 0.21, \emph{p} = 0.837; \emph{Figure \ref{fig:ioi-2}}). A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two condition (\emph{p} = 0.956).

@@

\hypertarget{key-overlap-time-kot-3}{%
\subsubsection{3.3.4. Key-Overlap Time (KOT)}\label{key-overlap-time-kot-3}}

>>>>

\abstract{
aaaa
}

@@

We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment. For data analysis, we excluded 5 participants due to experimental errors (3), an overall deviated tempo (2; outside 2 standard deviations from the average tempo across participants). Thirty-one participants (15 female) were included in data analysis. Most participants were right-handed (left; 2, both; 2). They had 12.45 years of practice on average (\emph{SD} = 5.66). All participants gave their informed consent before the experiment started and received vouchers for their participation. The study was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary.

@@

In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution. The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with notated articulation (\emph{p} = 0.021; \emph{Figure \ref{fig:ioi-1}}).

@@

Blah

@@

The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with notated dynamics (\emph{p} = 0.111; \emph{Figure \ref{fig:ioi-1}}).

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-1-1} \caption{\label{fig:kot-1}Experiment 1: KOT (ms) when playing the piece with either articulation (left) or dynamics (right). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box. Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-kot-1}
\end{figure}

@@

A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.972, \emph{SD} = 0.042{]} than in the performing condition {[}\emph{M} = 0.958, \emph{SD} = 0.034{]} while playing the piece with notated articulation (\emph{t}(18) = 2.64
, \emph{p} = 0.017; \emph{Figure \ref{fig:ioi-2}}). A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (\emph{p} = 0.026).

\hypertarget{key-overlap-ratio-kor}{%
\subsubsection{3.2.2. Key-Overlap Ratio (KOR)}\label{key-overlap-ratio-kor}}

Blah

@@

A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.962, \emph{SD} = 0.041{]} and the performing condition {[}\emph{M} = 0.961, \emph{SD} = 0.037{]} while playing the piece with notated articulation (\emph{t}(18) = 0.37, \emph{p} = 0.719; \emph{Figure \ref{fig:ioi-2}}). A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two condition (\emph{p} = 0.922).

@@

Blah

\hypertarget{key-overlap-ratio-kor-1}{%
\subsubsection{3.3.4. Key-Overlap Ratio (KOR)}\label{key-overlap-ratio-kor-1}}

Blah
